Proposed Solution

### 2017-01-22 CAV17/secs/arch.tex
2/3 of learner maybe

\para{Initial phase.}
We start with the assumption 
that we are given a number of (not necessarily correct) 
configuration files, called {\em training set}, 
belonging to the same system, such as MySQL or Apache. 
These files, therefore, follow similar patterns.
%In the following steps,
%we will exploit in a collection of learning algorithms 
%to build rules that describe a language model for the files.

\para{Translator.}
The translator module first parses the input training  
set of configuration files, and then transforms them into 
a more structured and typed intermediate representation.
When we infer the types of entries in a configuration file, 
the type of an entry cannot always be fully determined from 
a single value, since it is very hard to understand
the purposes of key-value entries in modern
software configuration files~\cite{xu15hey}.
We address this problem 
by introducing {\em probabilistic types}.
Rather than giving a variable a single type, 
we assign several types over a probability distribution. 
We can later use these more structured files
as a training set to learn the rules.

\para{Learning.}
The input of the learner is a set of files that have been translated
into well-structured intermediate representations. 
The learner employs a learning algorithm named {\em rule association 
algorithm} to generate various rules,
potentially used to handle different types of configuration errors,
including ordering errors, missing entry errors,
type errors and integer correlation errors.
These rules are the outputs of the learner, 
and will be used to detect errors later.
Because the translator gives the learner probabilistically typed entries,
the learner is also responsible for determining a type for each entry.

\para{Post Analysis.}
Finally, the logically structured representation of learned rules allows for a further post analysis stage.
This novel extension infers knowledge on the accepted rules that can improve the output of the system..
As a case study, we build a graph to model the learned rules.
We analyze the properties of this graph to construct an ordering of rules by their importance, 
  as well as to produce a measure of complexity for any configuration of the target system.
While the metrics in used in \app are effective, they are not intended to be exhaustive.
The information contained in the structured representation of the learned rules, 
  is a unique benefit of the learning algorithm, as contrasted with other machine learning techniques.

>>>>

\para{Initial phase}
We start with the assumption 
that we are given a number of (not necessarily correct) 
configuration files, called {\em training set}, 
belonging to the same system, such as MySQL or Apache. 
These files, therefore, follow similar patterns.
%In the following steps,
%we will exploit in a collection of learning algorithms 
%to build rules that describe a language model for the files.

\para{Translator}
The translator module first parses the input training  
set of configuration files, and then transforms them into 
a more structured and typed intermediate representation.
When we infer the types of entries in a configuration file, 
the type of an entry cannot always be fully determined from 
a single value, since it is very hard to understand
the purposes of key-value entries in modern
software configuration files~\cite{xu15hey}.
We address this problem 
by introducing {\em probabilistic types}.
Rather than giving a variable a single type, 
we assign several types over a probability distribution. 
We can later use these more structured files
as a training set to learn the rules.

\para{Learning}
The input of the learner is a set of files that have been translated
into well-structured intermediate representations. 
The learner employs a learning algorithm named {\em rule association 
algorithm} to generate various rules,
potentially used to handle different types of configuration errors,
including ordering errors, missing entry errors,
type errors and integer correlation errors.
These rules are the outputs of the learner, 
and will be used to detect errors later.
Because the translator gives the learner probabilistically typed entries,
the learner is also responsible for determining a type for each entry.

\para{Post Analysis}
Finally, the logically structured representation of learned rules allows for a further post analysis stage.
This novel extension infers knowledge on the accepted rules that can improve the output of the system.
As a case study, we build a graph to model the learned rules.
We analyze the properties of this graph to construct an ordering of rules by their importance, 
  as well as to produce a measure of complexity for any configuration of the target system.
While the metrics in used in \app are effective, they are not intended to be exhaustive.
The information contained in the structured representation of the learned rules, 
  is a unique benefit of the learning algorithm, as contrasted with other machine learning techniques.

### 2017-01-22 CAV17/secs/learner.tex
2/3 of learner maybe

The goal of the learner is to derive rules from the intermediate representation of the training set generated by the translator.
We an interface for uses to define rules that should be learned.
Each instance of the interface corresponds to a different class of of configuration errors, such as missing entry errors, ordering errors, and fine-grained value correlation errors. 
These errors can cause total system failures, but can also be more insidious, for example slowing down the system only when the server load increases beyond a certain threshold.

\para{Type}
In \app, probabilistic typing is implemented as an instance of the learning interface than can be easily extended to support more types. 
It

\para{Order}

\para{Missing}

\para{Integer Correlation}
Fine grained errors (Aaron)

>>>>

The goal of the learner is to derive rules from the intermediate representation of the training set generated by the translator.
We describe an interface to define the different classes of rules that should be learned.
Each instance of the interface corresponds to a different class of of configuration errors, such as missing entry errors, ordering errors, and integer correlation errors. 
These errors can cause total system failures, but can also be more insidious, for example slowing down the system only when the server load increases beyond a certain threshold.

A rule is an implication relationship, $X \implies p(X,Y)$, between two possibly empty sets of keywords $X,Y \subset \keys$, where $\keys$ is the set of unique keys from the training set and the predicate $p$ is the one of the classes of configuration errors (order, missing, etc).
Implicitly we interpret this to means that if the keywords $X\cup Y$ appear in a configuration file, we expect the predicate $p$ to hold.
The task of the learning algorithm is to transform a training set to a set of rules, weighted with \textit{support} and \textit{confidence}.
These two key metrics are taken from \textit{association rule learning}~\cite{agrawal1993mining}, a technique that can be summarized as inductive machine learning.
In the configuration verification domain, standard association rule learning is best suited to learn integer correlation and missing keyword rules.
In fact, a more specialized technique for learning ordering rules would be sequence mining~\cite{}, and again another approach may be a better fit for rules over a single keyword, as is needed for probabilistic types.
Since the full algorithm details are out-of-scope of this paper, here we only describe in detail the metrics which are most relevant to verification.

%TODO math def of support and confidence
Each power set of keywords, $\{X,Y\}$, is assigned a support and confidence measure during the learning process.
Support is the number of times the set of keywords in the proposed rule have been seen the in the training set.
Confidence is the number of times the rule predicate has held true over the given keywords.
In the learning process, these are given threshold, below which a rule will be reject for lack of evidence.

\para{Order}
Ordering errors take the form $X \implies order(X,Y)$, where $|X|,|Y|=1$ and the keyword $X$ must come before the keyword $Y$ in any configuration file.

\para{Missing}
$X \implies missing(X,Y)$, where $|X|,|Y|=1$ and the keyword $X$ must appear in the same file as the keyword $Y$ in any configuration file.

\para{Type}
The type rule is a set of rules over multiple predicates, which take the form $X \implies isType\ast(X)$, where $|X|=1, |Y|=0$ and $\ast$ matches all the basic types (string, int, etc).
In \app, probabilistic typing is implemented as an instance of the learning interface.
This module will specify the counting and resolution type judgments from Sec.~\ref{sec:ptypes}.
It

\para{Integer Correlation}
\app supports two types of integer correlation rules, coarse-grained and fine-grained.
Coarse-grained rules follow $X \implies compare(X,Y)$, where $|X|,|Y|=1$ and $compare \in \{<,=,>\}$, such that $X$ must hold $compare$ to $Y$.
Fine-grained rules follow $X \implies compare(X,Y)$, where $|X|=2,|Y|=1$ and $compare \in \{<,=,>\}$, such that for $k_1,\ k_2 \in X,\ k1*k2$ must hold $compare$ to $Y$.
These rules also implement a typing judgment over the keywords probabilistic type.
To avoid learning too many false positives, we restrict this rule to either $size*int=size$, $int*size=size$, or $int*int = int$.
Without probabilistic typing, we would also learn, for example, $int*int=size$.
%TODO Aaron

### 2017-01-22 CAV17/secs/limit.tex
2/3 of learner maybe

\para{Legal misconfigurations.}
Most of the existing proactive misconfiguration detection techniques, cannot handle configuration errors resulting from events that occurred during system runtime.
Such configuration errors are referred to as {\em legal misconfigurations}~\cite{yin11anempirical}. 
In particular, many parameter misconfigurations have perfectly legal types and values, but do not deliver the functionality intended by users. 
For example, after a website's traffic significantly increases, the parameter {\tt key\_buffer\_size} in MySQL may not  be able to handle increasingly more data traffic,
   thus leading to outage of the whole system.
While \app does not explicitly use any runtime information, we are still able to detect such errors, for example Sec.~\ref{sec:motiv4}.
Since \app relies on the collective knowledge of the training set, as long as the training set is mostly correct with respect to legal misconfigurations, these can still be detected.

\para{Misconfiguration across software components.}
As exposed by Yin {\em et al.}~\cite{yin11anempirical},
cross-software configuration correlation problems also account
for a considerable number of misconfiguration cases.
For example, in a LAMP-based Web server, one entry in the
PHP configuration file, {\tt mysql.max\_persistent = 400},
may make users encounter a ``too many connections'' error,
because a correlated entry in the underlying MySQL configuration
file assigns {\tt max\_connection} to 300, which is less
than the MySQL connection number in the PHP configuration file (\ie, 400).
It is quite difficult to detect such tricky errors
through learning approaches, because not only are users or engineers 
unaware of the hidden interactions~\cite{xu15systems},
but also it is hard to obtain global knowledge of the entire
configuration due to the business privacy concerns of
each software provider.
One possible solution to this problem might be to introduce
some cryptographic protocol, \eg, private set
intersection~\cite{kissner05privacy}, to privately extract the
overlapping entries, \eg, {\tt mysql.max\_connection} in the 
above MySQL and PHP case, for double-checking.

\para{Network configuration verification.}
\app mainly focuses on software configurations, \eg, MySQL and Apache,
so that our approach is limited to support network configuration
verification. This is because network configurations have quite
different representations, format, and rules from software configurations,
since network configurations are typically written in 
more domain-specific policy languages.
In fact, many network verification tools, 
\eg, NoD~\cite{lopes15checking} and 
Dobrescu {\em et al.}~\cite{dobrescu14software},
have been proposed to check whether network configurations
meet their specifications.

\para{More complex configuration structure.}
Currently, \app mainly targets key-value configuration files,
but in practice many systems, \eg, OpenStack~\cite{OpenStack},
employ very complex configuration format and structure,
which \app cannot handle.
Verifying such structurally complex configurations typically requires 
\app to learn a much more sophisticated language model,
which is challenging in practice.
Recent efforts~\cite{raychev15predicting, raychev16learning} 
may present possible solutions to this limitation.
These techniques can learn a call-graph from a training program set,
and check a new program based on properties extracted from this
generated call-graph. If we examine a structurally complex configuration 
file as a program with these tools, we may be able to use similar techniques 
to verify whether the configuration file violates a property of interest.

>>>>

%TODO move this - it is not a limitation
\para{Legal misconfigurations}
Most of the existing proactive misconfiguration detection techniques, cannot handle configuration errors resulting from events that occurred during system runtime.
Such configuration errors are referred to as {\em legal misconfigurations}~\cite{yin11anempirical}. 
In particular, many parameter misconfigurations have perfectly legal types and values, but do not deliver the functionality intended by users. 
For example, after a website's traffic significantly increases, the parameter {\tt key\_buffer\_size} in MySQL may not  be able to handle increasingly more data traffic,
   thus leading to outage of the whole system.
While \app does not explicitly use any runtime information, we are still able to detect such errors, for example Sec.~\ref{ex:fine}.
Since \app relies on the collective knowledge of the training set, as long as the training set is mostly correct with respect to legal misconfigurations, these can still be detected.

\para{Misconfiguration across software components}
As reported by Yin {\em et al.}~\cite{yin11anempirical},
cross-software configuration correlation problems also account
for a considerable number of misconfiguration cases.
For example, in a LAMP-based Web server, one entry in the
PHP configuration file, {\tt mysql.max\_persistent = 400},
may make users encounter a ``too many connections'' error,
because a correlated entry in the underlying MySQL configuration
file assigns {\tt max\_connection} to 300, which is less
than the MySQL connection number in the PHP configuration file (\ie, 400).
It is quite difficult to detect such tricky errors
through learning approaches, because not only are users or engineers 
unaware of the hidden interactions~\cite{xu15systems},
but also it is hard to obtain global knowledge of the entire
configuration due to the business privacy concerns of
each software provider.
One possible solution to this problem might be to introduce
some cryptographic protocol, \eg, private set
intersection~\cite{kissner05privacy}, to privately extract the
overlapping entries, \eg, {\tt mysql.max\_connection} in the 
above MySQL and PHP case, for double-checking.

\para{Network configuration verification}
\app mainly focuses on software configurations, \eg, MySQL and Apache,
so that our approach is limited to support network configuration
verification. This is because network configurations have quite
different representations, format, and rules from software configurations,
since network configurations are typically written in 
more domain-specific policy languages.
In fact, many network verification tools, 
\eg, NoD~\cite{lopes15checking} and 
Dobrescu {\em et al.}~\cite{dobrescu14software},
have been proposed to check whether network configurations
meet their specifications.

\para{More complex configuration structure}
Currently, \app mainly targets key-value configuration files,
but in practice many systems, \eg, OpenStack~\cite{OpenStack},
employ very complex configuration format and structure,
which \app cannot handle.
Verifying such structurally complex configurations typically requires 
\app to learn a much more sophisticated language model,
which is challenging in practice.
Recent efforts~\cite{raychev15predicting, raychev16learning} 
may present possible solutions to this limitation.
These techniques can learn a call-graph from a training program set,
and check a new program based on properties extracted from this
generated call-graph. If we examine a structurally complex configuration 
file as a program with these tools, we may be able to use similar techniques 
to verify whether the configuration file violates a property of interest.

### 2017-01-22 CAV17/secs/motiv.tex
2/3 of learner maybe

\para{Example~1: Ordering error.} 
Ordering errors were reported by Yin {\em et al.}~\cite{yin11anempirical} and our first
example illustrates how ordering errors can cause a system to crash. When a user configures PHP 
to run with the
Apache HTTP Server, most likely the user will take some already existing configuration files and adapt them
to suit her needs. The configuration file might contain, among others, 
the following lines:

@@

\para{Example~2: Fine-grained value correlation error.} 
Our second misconfiguration example~\cite{correlation} 
comes from a discussion on StackOverflow.
The user has configured her MySQL as in the following:

@@

\para{Example~3: Missing entry error.} 
Many critical system outages result from the fact that an important
entry was missing from the configuration file. 
We call such a problem a {\em missing entry error}.
In a public misconfiguration dataset~\cite{configdataset},
many MySQL failure reports were caused by
missing entry errors.
Below is a real-world missing entry error example~\cite{yin11anempirical}:
when a user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
achieved via the following entries in the OpenLDAP configuration file:

@@

\para{Example~4: Type errors.} 
Many system performance problems are caused by 
assigning incorrect type of values to some key in configuration
files. Consider the following real-world misconfiguration file: 
a user tries to install MySQL and she needs to initiate the path
of the log information generated by MySQL.
This user puts the following assignment in her MySQL
configuration file:

>>>>

\para{Example~1: Ordering error} 
Ordering errors were reported by Yin {\em et al.}~\cite{yin11anempirical} and our first
example illustrates how ordering errors can cause a system to crash. When a user configures PHP 
to run with the
Apache HTTP Server, most likely the user will take some already existing configuration files and adapt them
to suit her needs. The configuration file might contain, among others, 
the following lines:

@@

%TODO would be nice to have these as subsection or somehow be able to ref specific examples
\para{Example~2: Fine-grained value correlation error}
\label{ex:fine}
Our second misconfiguration example~\cite{correlation} 
comes from a discussion on StackOverflow.
The user has configured her MySQL as in the following:

@@

\para{Example~3: Missing entry error} 
Many critical system outages result from the fact that an important
entry was missing from the configuration file. 
We call such a problem a {\em missing entry error}.
In a public misconfiguration dataset~\cite{configdataset},
many MySQL failure reports were caused by
missing entry errors.
Below is a real-world missing entry error example~\cite{yin11anempirical}:
when a user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
achieved via the following entries in the OpenLDAP configuration file:

@@

\para{Example~4: Type errors} 
Many system performance problems are caused by 
assigning incorrect type of values to some key in configuration
files. Consider the following real-world misconfiguration file: 
a user tries to install MySQL and she needs to initiate the path
of the log information generated by MySQL.
This user puts the following assignment in her MySQL
configuration file:

### 2017-01-22 CAV17/secs/trans.tex
2/3 of learner maybe

The translator will convert each key-value assignment $k=v$ in the configuration file to a triple, $(k, v, \tau)$, where $\tau$ is the type of $v$. 
There are two major challenges in this step.
First is that configuration files' keywords are not necessarily unique and may have some additional context (modules or conditionals).
To solve this, we rely on the fact that keywords in a configuration file must be unique within their context, and rename all keywords with their context.
The keys for the sample configuration \ref{lst:1.cnf} would then be [``{\tt foo[server]}'',``{\tt bar[client]}''] internally to the intermediate representation.

@@

An additional challenge is that it is not always possible to fully determine the type of key based on a single example value. 
For this reason, we introduce \textit{probabilistic types}, as contrasted with \textit{basic types}.
In \app, the set of basic types contains strings, file paths, integers, sizes, and booleans. 
Taking the configuration \ref{lst:1.cnf}, we can assume {\tt foo} is a boolean type by the grammar of MySQL,
  but the keyword {\tt bar} could either be a integer type or a Boolean type.
Consider that we would like to learn a rule that the boolean keywords should have the same values ($eq(\texttt{foo},\texttt{bar})$).
We then have two potential situations; in the first case where {\tt bar}$::int$, we do not learn the above rule, but in the second case where {\tt bar}$::bool$, the rule is valid.
To resolve this ambiguity, instead of immediately assigning one type to a value, the translator assigns a distribution of types to a value based on examples from the learning set.
% an idea closely related to existentially quantified types~\cite{Launchbury93lazyfunctional}.

A probabilistic type is a set of counts over a set $\mathcal{T}$ of basic types.
Formally, we define a space of probabilistic types $\tilde{\mathcal{T}}$, where $\ptype=\{(\tau_1, c_1),\ldots,(\tau_n, c_n)\}$ is a set of pairs, such that $\ptype \in \tilde{\mathcal{T}}, \tau_i \in \mathcal{T}$, $c_i \in \mathbb{Z}$. 
Every keyword has a probabilistic type, expressed $k:\ptype$, which, without any evidence, remains {\scriptsize EMPTY}.
The count for $(\tau_i,c_i) \in \ptype$ is incremented each time a new example value that matches to type $\tau_i$ is encountered for a particular key, using the update notation $\update$.

>>>>

The translator will convert each key-value assignment $k=v$ in the configuration file to a triple, $(k, v, \tau)$, where $\tau$ is the type of $v$. 
There are two major challenges in this step.
First is that configuration files' keywords are not necessarily unique and may have some additional context (modules or conditionals).
To solve this, we rely on the fact that keywords in a configuration file must be unique within their context, and rename all keywords with their context.
The set of unique keys, $\keys$, for the sample training set in Figure ~\ref{fig:tset} would then be [``{\tt foo[server]}'',``{\tt bar[client]}''] internally to the intermediate representation.
We will call the set of key-value pairs $C$.

@@

\subsection{Probabilistic Types}
\label{sec:ptypes}
An additional challenge is that it is not always possible to fully determine the type of key based on a single example value. 
For this reason, we introduce \textit{probabilistic types}, as contrasted with \textit{basic types}.
In \app, the set of basic types contains strings, file paths, integers, sizes, and booleans. 
Taking the configuration \ref{lst:1.cnf}, we can assume {\tt foo} is a boolean type by the grammar of MySQL,
  but the keyword {\tt bar} could either be a integer type or a Boolean type.
Consider that we would like to learn a rule that the boolean keywords should have the same values ($eq(\texttt{foo},\texttt{bar})$).
We then have two potential situations; in the first case where {\tt bar}$::int$, we do not learn the above rule, but in the second case where {\tt bar}$::bool$, the rule is valid.
To resolve this ambiguity, instead of immediately assigning one type to a value, the translator assigns a distribution of types to a value based on examples from the learning set.
% an idea closely related to existentially quantified types~\cite{Launchbury93lazyfunctional}.

A probabilistic type is a set of counts over a set $\mathcal{T}$ of basic types.
Formally, we define a space of probabilistic types $\tilde{\mathcal{T}}$, where $\ptype=\{(\tau_1, c_1),\ldots,(\tau_n, c_n)\}$ is a set of pairs, such that $\ptype \in \tilde{\mathcal{T}}, \tau_i \in \mathcal{T}$, $c_i \in \mathbb{Z}$. 
Every keyword $k\in\keys$ has a probabilistic type, expressed $k:\ptype$, which, without any evidence, remains {\scriptsize EMPTY}.
The count for $(\tau_i,c_i) \in \ptype$ is incremented each time a new example value that matches to type $\tau_i$ is encountered for a particular key, using the update notation $\update$.
